\documentclass{sig-alternate-05-2015}


\begin{document}

% Copyright
%\setcopyright{acmcopyright}
% DOI
\doi{insert doi here}

% ISBN
\isbn{insert isbn here}

\title{BigBInf Micro Cloud Platform}

\numberofauthors{1}
\author{
% 1st. author
\alignauthor
Brendan D. Ball\\
       \affaddr{University of Cape Town}\\
}

\date{30 August 2015}


\maketitle
\begin{abstract}
[Abstract]
\end{abstract}
\keywords{Big Data; BioInformatics; Clouds}


\section{Background}
Cloud infrastructure would go a long way in simplifying the processing of big data. Collaboration between researchers can potentially improve if they are able to process the same raw data while being in different locations. This would be done by keeping the data in a cloud solution and allowing researchers access to execute code on the raw data. There has been an explorative push towards cloud solutions from Amazon, Google etc, but there are significant drawbacks. The raw data needs to be uploaded initially and researchers often resort to mailing hard drives \cite{baker2010next}. 
\\
\\
An example of progress towards a cloud solution specifically for bioinformatics is Cloud BioLinux, which is a community driven project focussed on next generation sequencing. It is a toolkit which makes it easy to deploy virtual machines with bioinformatics infrastructure to a cloud platform. It bundles specific packages used in next generation sequence analysis, thereby decreasing configuration time and increasing maintainability. Instances of Cloud BioLinux have been tested on the Amazon EC2 cloud platform and on a private Eucalyptus cloud \cite{krampis2012cloud}.
\\
\\
Micro clouds deployed on-site would overcome the challenge of uploading big data to a commercial cloud. However, since different research institutions would deploy their own micro clouds, a need for cloud interoperability arises to maintain the ability for researchers from different institutions to collaborate on the same data. The cloud interoperability will form a community cloud. A use case of a specific community cloud has some similar architectural properties to what we are looking for. These properties include autonomy (where each micro cloud will be managed independently), security, self management of nodes, and scalable \cite{jimenez2014deploying}.
\\
\\
The traditional approach to creating a cloud platform which allows users to run their own instances of operating systems (such as Amazon EC2) is using virtualisation technology. This includes both hardware level emulation support and the software needed to manage the virtualisation. These virtualisation schemes use machine level virtualisation \cite{fink2014docker}. A new method, known as containerization provides much of the same functionality with added benefits of lower resource usage and better performance. Containers are able to run native machine instructions compared to virtualisation emulating every machine instruction \cite{dua2014virtualization}. Of course this means that containers are only useful when complete virtualisation is not needed, instead containers allow isolated application deployment and portability.

\section{Design}
\subsection{Design Aims}
The main aim of the micro cloud platform is to allow code to be uploaded by a user and execute it on local data. Users will make use of the micro cloud by interacting with a front end web interface. The web interface should be easy to use and provide useful feedback on a job that a user has submitted. 
\\\\
Given that multiple jobs can be submitted by multiple people, these jobs have to be scheduled properly to allow successful completion of each job. The default scheduler will rely on a first-in-first-out (FIFO) queue to schedule jobs. The whole platform should be fairly modular to allow pluggable components. This will allow possibilities such as adding a more advanced scheduler or completely replacing the backend while keeping the front end web interface.
\\\\
The final aim is to form a community cloud using a centralised discovery service. This will improve scalability compared to a full mesh network between micro clouds. The community cloud will allow sharing of data and collaboration of users between multiple micro clouds.

\subsection{Constraints}
Jobs can only be run via docker containers. To create a job you have to create a Dockerfile.
\subsection{Approach}
Use Git for VCS
Unit testing for ensuring code correctness
Use of Kubernetes
Specifically designed for running docker containers in a cluster
Recently released version 1 (out of beta ready for production)
Easy to use and not cluttered like OpenStack
OpenStack solves a broader problem and wasn’t specifically designed for docker containers
Other container managers exist, but have to choose 1, Kubernetes looks most promising
\subsection{Evaluation}
Evaluate functionality by uploading and executing example code on given data
Evaluate user interface effectiveness based on feedback from a limited number of expert users
Possibly Michelle, Peter, Bagula

\subsection{Implementation}
Dockerfiles used for uploading code
Use Kubernetes as a docker manager
Storage support
Support OpenStack Swift storage API
Support S3 storage API
Web interface
backend implemented using python flask web framework

\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{ref} 


\end{document}
